---
title: "Class 8 Code"
author: "Nicky Wakim"
date: "2023-04-26"
output: html_document
---
<style type="text/css">
.main-container {
  max-width: 850px;
  margin-left: auto;
  margin-right: auto;
}
</style>

<style type="text/css">
body, td {
   font-size: 16px;
}
code.r{
  font-size: 16px;
}
pre {
  font-size: 16px
}
</style>

### Set up for the code
```{r}
library(tidyr)
library(dplyr)
library(epiDisplay)
library(kableExtra)
library(here)
load(here("data", "bc_diagnosis.Rda"))
```

## Simple Logistic Regression (continued from Class 7)

### Interpreting a continuous independant variable
```{r}
age_glm = glm(Late_stage_diag ~ Age, data = bc, 
               family = binomial())
summary(age_glm)
logistic.display(age_glm)
```

```{r}
library(gmodels)
age_10 = estimable(age_glm, c(0,10))
age_10
OR_age_10 = exp(c(OR = age_10$Estimate, 
                  L_CI = age_10$Estimate - 1.96*age_10$`Std. Error`, 
                  U_CI = age_10$Estimate + 1.96*age_10$`Std. Error`))
OR_age_10
```

## Multiple Logistic Regression
### Fitting the model
```{r}
multi_bc = glm(Late_stage_diag ~ Race_Ethnicity + Age, data = bc, 
               family = binomial())

summary(multi_bc)
```

### Wald test for single coefficients
```{r}
wald_test = summary(multi_bc)$coefficients
rownames(wald_test) = c("Intercept", "R_E: HL", "R_E: NH AIAN",
                        "R_E: NH API","R_E: NH B", "Age")
round(wald_test, 3)
```

### Likelihood ratio test
#### Example 1: Age
```{r message=FALSE, warning=FALSE}
multi_bc = glm(Late_stage_diag ~ Race_Ethnicity + Age, data = bc, 
               family = binomial())
r_e_bc = glm(Late_stage_diag ~ Race_Ethnicity, data = bc, 
               family = binomial())
library(lmtest)
lrtest(multi_bc, r_e_bc)
```

#### Example 2: Race and Ethnicity
```{r message=FALSE, warning=FALSE}
multi_bc = glm(Late_stage_diag ~ Race_Ethnicity + Age, data = bc, 
               family = binomial())
age_bc = glm(Late_stage_diag ~ Age, data = bc, 
               family = binomial())
library(lmtest)
lrtest(multi_bc, age_bc)
```

#### Example 3: Race and Ethnicity, and age!
```{r message=FALSE, warning=FALSE}
multi_bc = glm(Late_stage_diag ~ Race_Ethnicity + Age, data = bc, 
               family = binomial())
int_bc = glm(Late_stage_diag ~ 1, data = bc, 
               family = binomial())
library(lmtest)
lrtest(multi_bc, int_bc)
```

## Predicted/Estimated Probability
```{r}
newdata = with(bc, 
               data.frame(Race_Ethnicity = "NH Asian/Pacific Islander", 
                          Age = 61.70569))
pred = predict(multi_bc, newdata, se.fit = T, type="response")

LL_CI = pred$fit - qnorm(1-0.05/2) * pred$se.fit
UL_CI = pred$fit + qnorm(1-0.05/2) * pred$se.fit
c(Pred = pred$fit, LL = LL_CI, UL = UL_CI)
```

### Interpreting the coefficients
```{r}
multi_bc = glm(Late_stage_diag ~ Race_Ethnicity + Age, data = bc, 
               family = binomial())
exp(summary(multi_bc)$coefficients)
exp(confint.default(multi_bc))
```
