---
title: "Class 13 Code"
author: "Nicky Wakim"
date: "2023-05-21"
output: html_document
---
<style type="text/css">
.main-container {
  max-width: 1100px;
  margin-left: auto;
  margin-right: auto;
}
</style>

<style type="text/css">
body, td {
   font-size: 16px;
}
code.r{
  font-size: 16px;
}
pre {
  font-size: 16px
}
</style>

## Set up for the code
```{r message=FALSE, warning=FALSE}
library(car)
library(mctest)
library(epiDisplay)
library(kableExtra)
library(dplyr)
```

## Zero cell in contingency table

### One covariate
```{r}
ex1 = as.data.frame(rbind(cbind(outcome = rep(1, 39), x = c(rep("One", 7), rep("Two", 12), rep("Three", 20))), 
      cbind(outcome = rep(0, 21), x = c(rep("One", 13), rep("Two", 8))))) %>%
  mutate(outcome = as.numeric(outcome), 
         x = factor(x, levels = c("One", "Two", "Three")))
```

Running glm from the data
```{r}
ex1_glm = glm(outcome ~ x, data = ex1, family = binomial())
summary(ex1_glm)
logistic.display(ex1_glm)
```

Combine groups 2 and 3:
```{r}
ex1_23 = ex1 %>% mutate(x = factor(x, levels = c("One", "Two", "Three"), labels = c("One", "Two", "Two")))
ex1_23_glm = glm(outcome ~ x, data = ex1_23, family = binomial())
summary(ex1_23_glm)
```

Eliminate the third group:
```{r}
ex1_two = ex1 %>% filter(x != "Three")
ex1_two_glm = glm(outcome ~ x, data = ex1_two, family = binomial())
summary(ex1_two_glm)
```

Treat covariate as continuous:
```{r}
ex1_cont = ex1 %>% mutate(x = as.numeric(x))
table(ex1_cont$outcome, ex1_cont$x)

ex1_cont_glm = glm(outcome ~ x, data = ex1_cont, family = binomial())
summary(ex1_cont_glm)
logistic.display(ex1_cont_glm)
```

### Interaction
```{r}
outcome = c(rep(1, 35), rep(0, 25))
x1 = c(rep(1, 5), rep(0, 2), rep(1, 10), rep(0, 2), rep(1, 15), rep(0, 1),
       rep(1, 5), rep(0, 8), rep(1, 2), rep(0, 6), rep(0, 4))
x2 = c(rep(1, 7), rep(2, 12), rep(3, 16), 
       rep(1, 13), rep(2, 8), rep(3, 4))
ex2 = data.frame(outcome = outcome, x1 = x1, x2 = x2)

```

## Complete Separation
```{r}
y = c(0,0,0,0,1,1,1,1)
x1 = c(1,2,3,3,5,6,10,11)
x2 = c(3,2,-1,-1,2,4,1,0)
ex3 = data.frame(outcome = y, x1 = x1, x2= x2)
ex3
m1 = glm(y ~ x1 + x2, family=binomial)
summary(m1)

library(logistf)
m1_f = logistf(y ~ x1 + x2, family=binomial)
summary(m1_f)
```


## Collinearity

Bring in GLOW study again
```{r}
library(aplore3)
glow = glow500
glow2 = glow %>% mutate(raterisk2 = factor(raterisk, levels = c("Less", "Same", "Greater"), 
                                            labels = c("Less and Same", "Less and Same", "Greater"))) 
```

VIF
```{r}
main_eff = glm(fracture ~ age + height + priorfrac + momfrac + 
                 armassist + raterisk2, 
                 data = glow2, family = binomial)

glow_model = glm(fracture ~ age + height + priorfrac + momfrac + 
                   armassist + raterisk2 + weight + bmi, 
                 data = glow2, family = binomial)

library(car)
vif(main_eff)
vif(glow_model)
```


```{r}
glow3 = glow2 %>% mutate(height_sq = height^2)
height2 = glm(fracture ~ age + height + priorfrac + momfrac + 
                 armassist + raterisk2 + height_sq, 
                 data = glow3, family = binomial)
# summary(height2)
vif(height2)
```

Centering height helps with the VIF:
```{r}
glow4 = glow2 %>% mutate(height_c = (height - mean(height)), 
                         height_c_sq = height_c^2)
height_c_2 = glm(fracture ~ age + height_c + priorfrac + momfrac + 
                 armassist + raterisk2 + height_c_sq, 
                 data = glow4, family = binomial)
# summary(height_c_2)
vif(height_c_2)
```