---
title: "Class 15 Code"
author: "Nicky Wakim"
date: "2023-05-31"
output: html_document
---

```{=html}
<style type="text/css">
.main-container {
  max-width: 1100px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```
```{=html}
<style type="text/css">
body, td {
   font-size: 16px;
}
code.r{
  font-size: 16px;
}
pre {
  font-size: 16px
}
</style>
```
### Set up for the code

This includes the `LogisticDx` functions. The package `LogisticDx` was not installing on my computer, and as far as I know, not working for newer versions of R in general. I have emailed the maintainer of the package to see if they can update the package and get it working for newer versions of R. In the meantime, I have copied specific functions into the file `Logistic_Dx_Functions.R`. We will need to `source()` the R script so that we can use the functions. Please make sure you download the `Logistic_Dx_Functions.R` and include the needed file location to run it.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(epiDisplay)
library(kableExtra)
library(ggplot2)
library(aplore3)
library(GGally)
library(mfp)
library(ResourceSelection)
library(DescTools)
source("Logistic_Dx_Functions.R") # Edit this if your file is not in the same folder as this Rmd file
```

# GLOW study example

The GLOW study data is already a part of the `aplore3` package (stands for the "**Ap**plied **Lo**gistic **Re**gression **3**rd Edition).

```{r, fig.height=9, fig.width=9}
glow = glow500
```

# Overall Measures of Fit

## Pearson Residuals

```{r message=FALSE, warning=FALSE}
glow2 = glow %>% mutate(raterisk2 = factor(raterisk, levels = c("Less", "Same", "Greater"), 
                                            labels = c("Less and Same", "Less and Same", "Greater")), 
                        frac_num = as.numeric(fracture)-1) 
```

```{r message=FALSE, warning=FALSE, fig.keep = "none"}
cat_model = glm(frac_num ~ priorfrac + momfrac + 
                  armassist + raterisk2 + 
                  momfrac*armassist, 
                 data = glow2, family = "binomial")

gof(cat_model)$chiSq

hoslem.test(glow2$frac_num, fitted(cat_model), g = 10)
```

## AIC and BIC

```{r}
prelim_final = glm(fracture ~ age + height + priorfrac + 
                     momfrac + armassist + raterisk2 + 
                     age*priorfrac + momfrac*armassist, 
                 data = glow2, family = binomial)
```

```{r}
AIC(prelim_final)
BIC(prelim_final)

cat_model = glm(frac_num ~ priorfrac + momfrac + 
                  armassist + raterisk2 + 
                  momfrac*armassist, 
                 data = glow2, family = "binomial")
```

```{r}
AIC(cat_model)
BIC(cat_model)
```

# Logistic Regression Diagnostics

Fit the preliminary final model first. Then we use the `dx` function to extract the diagnostic values from the covariate patterns.
```{r}
prelim_final = glm(fracture ~ age + height + priorfrac + 
                     momfrac + armassist + raterisk2 + 
                     age*priorfrac + momfrac*armassist, 
                 data = glow2, family = binomial)
```

```{r}
diagnost_pf = dx(prelim_final)
head(diagnost_pf)
dim(diagnost_pf)
```

Plot the change in standardized Pearson residuals against the estimated/predicted probabilities. Remember, this is the change in the 
```{r fig.height=5, fig.width=6}
ggplot(diagnost_pf) + geom_point(aes(x=P, y=dChisq)) + 
  xlab("Estimated/Predicted Probability of Fracture") +
  ylab("Change in Std. Pearson Residual") +
  theme(text = element_text(size = 16))
```

Plot the change in standardized deviance residuals against the estimated/predicted probabilities. 
```{r fig.height=5, fig.width=6}
ggplot(diagnost_pf) + geom_point(aes(x=P, y=dDev)) + 
  xlab("Estimated/Predicted Probability of Fracture") +
  ylab("Change in Std. Deviance Residual") +
  theme(text = element_text(size = 16))
```

Plot the change in coefficient estimates against the estimated/predicted probabilities. 
```{r fig.height=5, fig.width=6}
ggplot(diagnost_pf) + geom_point(aes(x=P, y=dBhat)) + 
  xlab("Estimated/Predicted Probability of Fracture") +
  ylab("Change in Coefficient Estimates") +
  theme(text = element_text(size = 16))
```

```{r fig.height=5, fig.width=6}
ggplot(diagnost_pf) + geom_point(aes(x=P, y=h)) + 
  xlab("Estimated/Predicted Probability of Fracture") +
  ylab("Change in Coefficient Estimates") +
  theme(text = element_text(size = 16))
```

```{r}
diagnost_points = diagnost_pf %>% 
  mutate(Cov_patt = 1:nrow(.)) %>%
  filter((sPr)^2 > 12.5 | (sdr)^2 > 4 | dBhat > 0.125) %>%
  dplyr::select(Cov_patt, y, P, h, dChisq, dDev, dBhat) %>%
  round(., 2)
kable(diagnost_points) %>% kable_styling(bootstrap_options = c("striped", "hover"), full_width = F)
```

```{r}
diagnost_points2 = diagnost_pf %>% 
  mutate(Cov_patt = 1:nrow(.)) %>%
  filter(Cov_patt %in% c(391, 428, 436, 454)) %>%
  round(., 2)
glow3 = glow2 %>%
  filter(!(age == 65 & height == 167 & priorfrac == "No"))
```
