---
title: "Class 14 Code"
author: "Nicky Wakim"
date: "2023-05-24"
output: html_document
---
<style type="text/css">
.main-container {
  max-width: 1100px;
  margin-left: auto;
  margin-right: auto;
}
</style>

<style type="text/css">
body, td {
   font-size: 16px;
}
code.r{
  font-size: 16px;
}
pre {
  font-size: 16px
}
</style>

### Set up for the code
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(epiDisplay)
library(kableExtra)
library(ggplot2)
library(aplore3)
library(GGally)
library(mfp)
source("~/Library/CloudStorage/OneDrive-OregonHealth&ScienceUniversity/Teaching/S2023_BSTA_513_613/Lectures/Week_9_Model_fit/Class_15/Code/Logistic_Dx_Functions.R")
```

## GLOW study example
The GLOW study data is already a part of the `aplore3` package (stands for the "**Ap**plied **Lo**gistic **Re**gression **3**rd Edition).
```{r, fig.height=9, fig.width=9}
glow = glow500
```

## Pearson Residuals
```{r message=FALSE, warning=FALSE, fig.keep = "none"}
glow2 = glow %>% mutate(raterisk2 = factor(raterisk, levels = c("Less", "Same", "Greater"), 
                                            labels = c("Less and Same", "Less and Same", "Greater")), 
                        fracture1 = as.numeric(fracture)-1) 
cat_model = glm(fracture1 ~ priorfrac + momfrac + armassist + raterisk2 + momfrac*armassist, 
                 data = glow2, family = "binomial")
summary(cat_model)

gof(cat_model)$chiSq

## Still can't find good replacement for LogisticDx package
```

## Hosmer-Lemeshow Test
```{r}
glow2 = glow %>% mutate(raterisk2 = factor(raterisk, levels = c("Less", "Same", "Greater"), 
                                            labels = c("Less and Same", "Less and Same", "Greater"))) 
```

```{r}
prelim_final = glm(fracture ~ age + height + priorfrac + momfrac + armassist + raterisk2 +
              age*priorfrac + momfrac*armassist, 
                 data = glow2, family = binomial)
summary(prelim_final)
```

Need to make sure that the observed outcome, fracture, is in a numeric scale (so takes 0 or 1 values). In the above code, I am using the outcome as a factor, so I will change it here.
```{r}
glow2 = glow2 %>% mutate(frac_num = as.numeric(fracture)-1)

library(ResourceSelection)
hoslem.test(glow2$frac_num, fitted(prelim_final), g = 10)
```

## ROC Curve and AUC
```{r fig.height=5, fig.width=6, message=FALSE, warning=FALSE}
library(ggplot2)
library(pROC)


predicted <- predict(prelim_final, glow2, type="response")

#define object to plot and calculate AUC
rocobj <- roc(glow2$fracture, predicted)
auc <- round(auc(glow2$fracture, predicted),4)

#create ROC plot
ggroc(rocobj, colour = 'steelblue', size = 2) +
  ggtitle(paste0('ROC Curve ', '(AUC = ', auc, ')')) +
  theme(text = element_text(size = 16)) +
  xlab("False Positive Rate (1 - Specificity)") +
  ylab("True Positive Rate (Sensitivity)")
```

## AIC and BIC
```{r}
prelim_final = glm(fracture ~ age + height + priorfrac + 
                     momfrac + armassist + raterisk2 + 
                     age*priorfrac + momfrac*armassist, 
                 data = glow2, family = binomial)
```
```{r}
AIC(prelim_final)
BIC(prelim_final)

prelim_final2 = glm(fracture ~ age + height + priorfrac + 
                      momfrac + armassist + raterisk2 + 
                      age*height, 
                 data = glow2, family = binomial)
```

```{r}
AIC(prelim_final2)
BIC(prelim_final2)
```

